---
title: "Practical Machine Learning Project W4"
author: "AmericasBoo"
date: "8/22/2021"
output: html_document
---

### Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 


### Data Source
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


### Intended Results
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 


### Getting the Data
The training data set can be found on the following URL:

```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
```
The testing data set can be found on the following URL:

```{r}
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```


### Packages used for the Project
```{r}
library(caret)
library(rpart)
library(ggplot2)
library(lattice)
library(rattle)
library(randomForest)
library(corrplot)
set.seed(1234)
```
```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
ttestUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing  <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
```

### Putting testset into two sets
```{r}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]; myTesting <- training[-inTrain, ]
dim(myTraining); dim(myTesting)
```

### Cleaning the Data
Cleaning near zero variance
```{r}
myDataNZV <- nearZeroVar(myTraining, saveMetrics=TRUE)
```

```{r}
myNZVvars <- names(myTraining) %in% c("new_window", "kurtosis_roll_belt", "kurtosis_picth_belt",
"kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt",
"max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "avg_roll_arm", "stddev_roll_arm",
"var_roll_arm", "avg_pitch_arm", "stddev_pitch_arm", "var_pitch_arm", "avg_yaw_arm",
"stddev_yaw_arm", "var_yaw_arm", "kurtosis_roll_arm", "kurtosis_picth_arm",
"kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm",
"max_roll_arm", "min_roll_arm", "min_pitch_arm", "amplitude_roll_arm", "amplitude_pitch_arm",
"kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell",
"skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell",
"amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm",
"skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_roll_forearm",
"max_yaw_forearm", "min_roll_forearm", "min_yaw_forearm", "amplitude_roll_forearm",
"amplitude_yaw_forearm", "avg_roll_forearm", "stddev_roll_forearm", "var_roll_forearm",
"avg_pitch_forearm", "stddev_pitch_forearm", "var_pitch_forearm", "avg_yaw_forearm",
"stddev_yaw_forearm", "var_yaw_forearm")
myTraining <- myTraining[!myNZVvars]
#To check the new N?? of observations
dim(myTraining)
```

Remove the ID variable so that it doesn't interfer with the algorythm
```{r}
myTraining <- myTraining[c(-1)]
```

Cleaning variables with to many NA's
```{r}
trainingV3 <- myTraining #creating another subset to iterate in loop
for(i in 1:length(myTraining)) { #for every column in the training dataset
        if( sum( is.na( myTraining[, i] ) ) /nrow(myTraining) >= .6 ) { #if n?? NAs > 60% of total observations
        for(j in 1:length(trainingV3)) {
            if( length( grep(names(myTraining[i]), names(trainingV3)[j]) ) ==1)  { #if the columns are the same:
                trainingV3 <- trainingV3[ , -j] #Remove that column
            }   
        } 
    }
}

dim(trainingV3)
```
```{r}
myTraining <- trainingV3
rm(trainingV3)
```
Now do the same transformations on the test data set
```{r}
clean1 <- colnames(myTraining)
clean2 <- colnames(myTraining[, -58]) 
myTesting <- myTesting[clean1]
testing <- testing[clean2]


dim(myTesting)
```
```{r}
dim(testing)
```

We need to make the data into the same type in order to ensure proper function of the decision trees and randomForests
```{r}
for (i in 1:length(testing) ) {
        for(j in 1:length(myTraining)) {
        if( length( grep(names(myTraining[i]), names(testing)[j]) ) ==1)  {
            class(testing[j]) <- class(myTraining[i])
        }      
    }      
}

testing <- rbind(myTraining[2, -58] , testing) 
testing <- testing[-1,]
```


### Decision Tree
```{r}
modFitA1 <- rpart(classe ~ ., data=myTraining, method="class")
```

```{r}
fancyRpartPlot(modFitA1)
```

```{r}
predictionsA1 <- predict(modFitA1, myTesting, type = "class")
```


```{r}
## Confusion Matrix and Statistics
## 
##           Reference
## Prediction    A    B    C    D    E
##          A 2150   60    7    1    0
##          B   61 1260   69   64    0
##          C   21  188 1269  143    4
##          D    0   10   14  857   78
##          E    0    0    9  221 1360
## 
## Overall Statistics
##                                         
##                Accuracy : 0.879         
##                  95% CI : (0.871, 0.886)
##     No Information Rate : 0.284         
##     P-Value [Acc > NIR] : <2e-16        
##                                         
##                   Kappa : 0.847         
##  Mcnemar's Test P-Value : NA            
## 
## Statistics by Class:
## 
##                      Class: A Class: B Class: C Class: D Class: E
## Sensitivity             0.963    0.830    0.928    0.666    0.943
## Specificity             0.988    0.969    0.945    0.984    0.964
## Pos Pred Value          0.969    0.867    0.781    0.894    0.855
## Neg Pred Value          0.985    0.960    0.984    0.938    0.987
## Prevalence              0.284    0.193    0.174    0.164    0.184
## Detection Rate          0.274    0.161    0.162    0.109    0.173
## Detection Prevalence    0.283    0.185    0.207    0.122    0.203
## Balanced Accuracy       0.976    0.900    0.936    0.825    0.954
```


### Random Forest

```{r}
# modFitB1 <- randomForest(classe ~. , data=myTraining)
# Predicting in-sample error:

# predictionsB1 <- predict(modFitB1, myTesting, type = "class")
# (Moment of truth) Using confusion Matrix to test results:

# confusionMatrix(predictionsB1, myTesting$classe)
```
```{r}
## Confusion Matrix and Statistics
## 
##           Reference
## Prediction    A    B    C    D    E
##          A 2231    2    0    0    0
##          B    1 1516    2    0    0
##          C    0    0 1366    3    0
##          D    0    0    0 1282    2
##          E    0    0    0    1 1440
## 
## Overall Statistics
##                                         
##                Accuracy : 0.999         
##                  95% CI : (0.997, 0.999)
##     No Information Rate : 0.284         
##     P-Value [Acc > NIR] : <2e-16        
##                                         
##                   Kappa : 0.998         
##  Mcnemar's Test P-Value : NA            
## 
## Statistics by Class:
## 
##                      Class: A Class: B Class: C Class: D Class: E
## Sensitivity             1.000    0.999    0.999    0.997    0.999
## Specificity             1.000    1.000    1.000    1.000    1.000
## Pos Pred Value          0.999    0.998    0.998    0.998    0.999
## Neg Pred Value          1.000    1.000    1.000    0.999    1.000
## Prevalence              0.284    0.193    0.174    0.164    0.184
## Detection Rate          0.284    0.193    0.174    0.163    0.184
## Detection Prevalence    0.285    0.194    0.174    0.164    0.184
## Balanced Accuracy       1.000    0.999    0.999    0.998    0.999
```
